---
title: "Chapter 12 Exercises: Normal Distribution"
output:
  github_document:
    fig_width: 7
    fig_height: 4
---

```{r load packages,include = FALSE}
set.seed(1431)
library(ggplot2)
library(dplyr)
```

**Question #1:**
What is the probability of observing a value five sigma greater than the mean or more?  

**Answer #1:**
We can solve this by generating a range of values with mean = 0 and sd = 1.  From this, we can integrate from 5 SDs to infinity.  

```{r calc q1}
data <- data.frame(value = rnorm(500000,mean=0,sd=1))
ggplot(data,aes(x=value)) + geom_density() + geom_vline(xintercept=5,linetype='dashed',colour='red') + xlab('Temperature') + ylab('Density')

integrate(function(x) dnorm(x,mean=0,sd=1),5,1000)
```  

By simply looking at the distribution above, we can see that there is an extremely low probability of a value greater than 5 SDs from the mean.  It is clear this value is nearly zero.  But how close to zero is this value?  We can calculate this by integrating from 5 to positive infinity.  I've done this above and we can see that the probability is around .00000008%.  

* * *
* * *  

**Question #2:**
A fever is any temperature greater than 100.4 degrees fahrenheit.  Given the following measurements, what is the probability that the patient has a fever?  

100, 99.8, 101, 100.5, 99.7  

**Answer #2:**
To begin, lets visualize the distribution we are examining.  In the code below, I calculate the mean and the standard deviation of the values we are given and generate 500k points from a normal distribution with these parameters.   

```{r q2 calc}
temp <- c(100, 99.8, 101, 100.5, 99.7)

mean = mean(temp)
sd = sd(temp)

data <- data.frame(temp = rnorm(500000,mean=mean,sd=sd))
ggplot(data,aes(x=temp)) + geom_density() + geom_vline(xintercept=100.4,linetype='dashed',colour='red') + xlab('Temperature') + ylab('Density')

max_value = max(data$temp)
```

Above, I have marked the point we are interested in (temp = 100.4).  To answer our question, we want to sum the area under the curve to the right of this dashed line.  This is essentially calculating the probability that our value is 100.4 or greater.  

```{r q2b calc}
integrate(function(x) dnorm(x,mean=mean,sd=sd),100.4,max_value)
```

From the above, we conclude that there is about a 35.6% probability that our patient has a fever.  

* * *
* * *  

**Question #3:**  
Suppose in Chapter 11 we tried to measure the depth of a well by timing coin drops and got the following values:  2.5, 3, 3.5, 4, 2  

The distance an object falls can be calculated (in meters) with the following formula:  distance = 1/2 x G x time^2  

Where G is 9.8 m/s/s.  What is the probability that the well is over 500 meters deep?  

**Answer #3:**  
To begin, lets again create a vector with our values (transformed from time to distance).  Let us then visualize the distribution.  

```{r calc q3}
time <- c(2.5, 3, 3.5, 4, 2)

mean = mean(time)
sd = sd(time)

data <- data.frame(time = rnorm(500000,mean=mean,sd=sd))
ggplot(data,aes(x=time)) + geom_density() + xlab('Seconds') + ylab('Density')
```  

That will work.  Now we will want to figure out how many seconds is ~500 meters.  Using the formula above, we can write a function that will solve this for us.  Once we have the value we need, we can integrate from that value to infinity across the distribution.   

```{r function for q3}
meters_to_seconds <- function(value) {
x <- seq(1,10000,.05)
meters = (1/2)*(9.8)*(x^2)

data <- data.frame(x,meters,solve=value)
data <- data %>% mutate(diff = abs(meters-solve)) %>% mutate(ranker = rank(diff)) %>% filter(ranker == 1)
seconds <- data$x
rm(data)

return(seconds)
}
seconds <- meters_to_seconds(500)
seconds

ggplot(data,aes(x=time)) + geom_density() + geom_vline(xintercept=seconds,linetype='dashed',colour='red') + xlab('Seconds') + ylab('Density')

integrate(function(x) dnorm(x,mean=mean,sd=sd),seconds,1000)  
## note that 1000 is somewhat arbitrary here.  It is simply a rough substitute for` 'infinity'
```  

From the above, we can see that it is nearly impossible that the well is 500 meters deep.  


* * *  
* * *  

**Question #4:**  
What is the probability that there is no well? By 'no well', we mean that the well is ~0 meters deep.  

**Answer #4:**  
First, let us calculate how much time it would take to drop around 0 meters.  

```{r q4 calc time}
seconds <- meters_to_seconds(0)
ggplot(data,aes(x=time)) + geom_density() + geom_vline(xintercept=seconds,linetype='dashed',colour='red') + xlab('Seconds') + ylab('Density')

integrate(function(x) dnorm(x,mean=mean,sd=sd),-1000,seconds)  
```  

Theoretically, it would take somewhere between 0 and 1 second to drop a distance of 0 meters.  If we integrate from negative infinity to one, we find that the probability that the well is 0 meters is around 0.5%.  This seems implausible and is mostly a function of our poor estimate for how many seconds it would take a coin to drop 0 meters.  Realistically, this number is closer to zero than one, but we integrate across this entire space.  

As we lower our limits for the integration from 1 to zero, we would find that this probability gets closer and closer to zero (although it technically never reaches zero).  
