---
title: "Chapter 13 Exercises: The PDF, CDF, and Quantile Function"
output:
  github_document:
    fig_width: 7
    fig_height: 4
---

```{r load packages,include = FALSE}
set.seed(1431)
library(ggplot2)
library(dplyr)
```

**Question #1:**
Using the code example for plotting the PDF on page 127, plot the CDF and quantile functions.  

**Answer #1:**
We can solve this by generating a range of values with mean = 0 and sd = 1.  From this, we can integrate from 5 SDs to infinity.

```{r calc q1}
alpha = 300
beta = 40000-alpha

xs <- seq(0.005,0.01,by=.00001)
density <- dbeta(xs,alpha,beta)

data <- data.frame(xs,density)

ggplot(data,aes(x=xs)) + stat_ecdf()

xs <- seq(.005,.01,by=0.00001)
plot(xs,dbeta(xs,alpha,beta),type='l',lwd=3)
```

By simply looking at the distribution above, we can see that there is an extremely low probability of a value greater than 5 SDs from the mean.  It is clear this value is nearly zero.  But how close to zero is this value?  We can calculate this by integrating from 5 to positive infinity.  I've done this above and we can see that the probability is around .00000008%.

* * *
* * *

**Question #2:**  
Returning to the task of measuring snowfall from Chapter 10, say you have the following measurements (in inches) of snowfall:  

7.8,9.4,10.0,7.9,9.4,7.0,7.0,7.1,8.9,7.4  

What is your 99.9 percent CI for the true value of snowfall?  

**Answer #2:**  
To answer this, we need to first calculate the mean and the standard deviation from the measurements for snowfall.  

```{r q2 p1 calc}
x <- c(7.8, 9.4, 10.0, 7.9, 9.4, 7.0, 7.0, 7.1, 8.9, 7.4)
mean <- mean(x)
sd <- sd(x)
```

We will then define the lower and upper bound of our CIs for a 99.9% interval.  Because our distribution is symmetrical, we can take an ever share of (1-.999) from both sides of the distribution.  This can be defined in R as:  

```{r q2 p2 calc}
bounds <- (1-.999)/2
lower <- bounds
upper <- 1-bounds
```

We can now use the qnorm function to find the value of the number we are looking for at each end of our distribution.  

```{r q2 p3 calc}
ci_lower <- qnorm(lower,mean=mean,sd=sd)
ci_upper <- qnorm(upper,mean=mean,sd=sd)
ci <- c(ci_lower,ci_upper)

ci
```

And there we have it!  Lets finally visualize our 99.9% CI.  

```{r q2 p4 calc}
data <- data.frame(value = rnorm(1000000,mean=mean,sd=sd))

ggplot(data,aes(x=value)) + geom_density() + geom_vline(xintercept=ci_lower,linetype='dashed',colour='red') + geom_vline(xintercept=ci_upper,linetype='dashed',colour='red') + xlab('inches')
```
