---
title: "Chapter 12 Exercises: Normal Distribution"
output:
  github_document:
    fig_width: 7
    fig_height: 4
---

```{r load packages,include = FALSE}
set.seed(1431)
library(ggplot2)
library(dplyr)
```

**Question #1:**
What is the probability of observing a value five sigma greater than the mean or more?

**Answer #1:**
We can solve this by generating a range of values with mean = 0 and sd = 1.  From this, we can integrate from 5 SDs to infinity.

```{r calc q1}
data <- data.frame(value = rnorm(500000,mean=0,sd=1))
ggplot(data,aes(x=value)) + geom_density() + geom_vline(xintercept=5,linetype='dashed',colour='red') + xlab('Temperature') + ylab('Density')

integrate(function(x) dnorm(x,mean=0,sd=1),5,1000)
```

By simply looking at the distribution above, we can see that there is an extremely low probability of a value greater than 5 SDs from the mean.  It is clear this value is nearly zero.  But how close to zero is this value?  We can calculate this by integrating from 5 to positive infinity.  I've done this above and we can see that the probability is around .00000008%.

* * *
* * *
