---
title: "Chapter 5 Exercises: The Beta Distribution"
output:
  github_document:
    fig_width: 7
    fig_height: 4
---

**Question #1:**  
You want to use the beta distribution to determine whether or not a coin you have is a fair coin-meaning that the coin gives you heads and tails equally.  You flip the coin 10 times and get 4 heads and 6 tails.  Using the beta distribution, what is the probability that the coin will land on heads more than 60% of the time?    

**Answer #1:**  
First, let us get a visual of the beta distribution with these parameters (beta(4,6)).  

```{r q1 calc}
p = seq(0,1,length=100)
plot(p,dbeta(p,4,6))
abline(v=.6)
```  

To answer our question, we are strictly interested in summing the area under the curve to the right of 0.6 in the plot above, as designated above with the vertical line.  We can do this by using the integrate function in R.  

```{r q1b calc}
integrate(function(x) dbeta(x,4,6),.6,1)
```  

From the above, we conclude that there is about a 10% chance that the true probability of getting heads is 60% or greater given our observed data.

* * *  
* * *  

**Question #2:**  
You flip the coin 10 more times and now have 9 heads and 11 tails total.  What is the probability that the coin is fair, using our definition of fair, give or take 5 percent?  

**Answer #2:**  
We will use a similar approach to above with a few key changes.  Namely, we will update our beta distribution parameters as well as change the interval from which we want to integrate to calculate our probability.  Lets start by visualizing our beta distribution and marking the area we want to integrate across.  

```{r q2 calc}
plot(p,dbeta(p,9,11))
abline(v=.45) ## 50/50 split - 5%
abline(v=.55) ## 50/50 split + 5%

integrate(function(x) dbeta(x,9,11),.45,.55)
```  

From the above, we can see that there is about a 30% chance of the coin's fairness (+/- 5% from 50%).  

* * *  
* * *  

**Question #3:**  
Data is the best way to become more confident in your assertions.  You flip the coin 200 more times and end up with 109 heads and 111 tails.  Now what is the probability that the coin is fair, give or take 5%?  

**Answer #4:**  
To answer this question, we will simply update the code used in Q3 with the new values for alpha and beta.  

```{r q3 calc}
plot(p,dbeta(p,109,111))
abline(v=.45) ## 50/50 split - 5%
abline(v=.55) ## 50/50 split + 5%

integrate(function(x) dbeta(x,109,111),.45,.55)
```  

From the above, we can see that there is about a ~86% chance of the coin's fairness (+/- 5% from 50%).  
